Functional oxide layers for electrical isolation and chemical passivation of steel substrates

### 2022-08-29 Thesis/chapter/03-theoretical-background.tex
pyDOE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Machine Learning and Statistics}
\ds{
introduction 
    why I chose this practicum, 
        bcs wanted to check if lab is for me 
    why i chose to include ml 
        bcs I wanted what i learned during undergrad courses help me find a good solution and 
        wanted to dive deeper into the ml field as it is very interesting for 
what is ML? 
what kind of ML techniques? 
statistics:
what is the difference between ml and stat? 
    superficially very similar but different goals and diff rahmenbedingungen 
DOE is a very old, much information less experiments 
    how to design experiment: already many insights for next time 
statistics: 
    help me find insight from data 
    ANOVA 
    linear regression 
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Why introduce machine learning to this project? 
I wanted to apply and delve into what I've been studying during my undergrad courses. 
This seemed like the perfect oppurtunity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Artificial Inteligence and Machine Learning}
%How can \gls{ai} and \gls{ml} be defined? 
The history of \gls{ai} goes back to the middle of the \nth{20} century. 
Researchers from the emerging field came together at the Dartmouth conference and the term "\gls{ai}" was coined\cite{McCarthy1955}. 
\Gls{ai}'s  history is beautifully depicted by McCorducks' 1982 book "Machines Who Think"\cite{McCorduck1982,Apter1982}, which focuses on the great minds behind the advances.
%For a more in depth dive into the history of AI consult the beautifully written review\cite{Apter1982} of McCorduck's 1982 book "Machines Who Think"\cite{McCorduck1982}, which focuses also on the great minds behind advances of \gls{ai}.
Pioneers like Alan Turing thought a lot about how to define, test and implement \gls{ai}\cite{Howard2019}. 
One example how to measure \gls{ai} is to let it play chess against a human\cite{Silver2017} 
(in 1997 a chess computer called Deep Blue won against the World Chess Champion Garry Kasparov for the first time\cite{Feng1999}).
Another test ingenioused by Alan Turing, is that a human communicates with an unknown entity (in written form) and must judge if they are dealing with a machine or a human being. 
It is easy for humans to come up with question to detect an AI, 
but when reading AI written articles\cite{gpt2020}, it's easy to see this test being passed in the near future.

@@

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Machine Learning Methods}
%
\Gls{ml} is at the base of most \gls{ai}s.
It is a umbrella term for programs with instructions to learn from data, i.e. gain knowledge, categorise, predict and make decisions based on data. 
%
There is a platitude of different machine learning methods: 
they can be divided into supervised (training set is labeled) and unsupervised (exploratory).  
An orthogonal division can be made by regression (continuous data) versus classification (discrete, categorical data). 
Independent from these $2\times2$ categories there are multiple ways to let machines learn from data.
\Gls{nn} (one of the most popular architectures for big data\cite{Chiroma2019}) are loosely modeled after the brain\cite{bishop1994neural}.
%
Artificial neurons (also called nodes), which are arranged in layers, 
are connected to each of the neurons of previous and next layers
and the weights (parameters of intensity), with which the data is routed from one neuron to another, 
are optimised during training. 
%
Convolutional-layer \gls{nn} excel in picture recognition\cite{Lecun1995conv} and in quantum mechanics too\cite{westermayr2020combining}.
Other common architectures include linear regression, kernel ridge regression, support vector regression.
There are also lesser know algorithms like 
evolutionary algorithms (e.g. \gls{ga} and \gls{pso}).
%
One advantage of evolutionary algorithms is that they start with a small data set
and periodically request new data in order to solve the problem iteratively.
\td{This is exactly what I needed because every experiment is very time intensive.}

@@

\Gls{pso} also uses a starting population of particles where each experiment (particle) 
is represented by its independent and dependent variables. 
It was originally inspired by the behaviour of bird flocks and fish schools.
Each particle has an associated position and velocity. 
Every movement across the search space is stochastic and influenced by its particle velocity and position as well as its and the swarm's best known position.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Design of Experiment} %DOE
%\epigraph{"The real purpose of experiment design is to maximize the information content of the data within the limits imposed by the given constraints."}{Grahem C. Goodwin\cite{goodwin1977experiment}}
%
\begin{quote}
	{"The real purpose of experiment design is to maximize the information content of the data within the limits imposed by the given constraints."}
	- {Grahem C. Goodwin\cite{goodwin1977experiment}}
\end{quote}
%
In two cases a deliberate \gls{doe} is beneficial.
If the query of a new data point is very expensive, it is favourable to actively chose the query (e.g. drilling for oil or quantum chemical calculations). 
If the query space is so vast, that randomly querying might explore domains which might lead to uninteresting or even misleading information.
% 
At the beginning of an experiment its constraints must be determined. 
%
Constraints for a given experiment might include range of input and output variables as well as total time available and total number of samples/experiments that can be taken.\cite{goodwin1977experiment}
%

A naive approach to an experiment design is the full factorial \gls{doe}.
Each possible combination of discrete values is tested. 
While this is the most informative design it will more often than not be infeasible due the curse of dimensionality\cite{cherkassky1998learning}.
The 2-level factorial design provides an alternative with $2^d$ (where $d$ is the number of independent variables) experiments. 
Drawbacks of 2-level factorial include no data about the inside of the search space and infeasibility for moderate dimensional problems.
In a full factorial or 2-level factorial design most experiments are redundant and most resources will be spent exploring high-order interaction effects\cite{gunst2009fractional}.
In order to overcome these obstacles a certain number of experiments can be chosen randomly from the search space. 
%A straight forward design can be to randomly chose a certain number of experiments from the search space.

The Plackett-Burman\cite{vanaja2007design,miller2001using,wang1995hidden} design is a special case of 2-level fractional factorial design, 
where the number of needed experiments is smaller than $d+2$ and 
each combination of levels for any pair of factors appears the same number of times. 
A drawback of randomization and 2-level factorial designs is that the sample set is likely not evenly distributed across the search space\cite{viana2016tutorial}. 
The Hammersley design\cite{viana2016tutorial,diwekar1997efficient} is based on the Hammersley sequence and produces space filling data points. 
The Latin hypercube \gls{doe}\cite{viana2016tutorial,diwekar1997efficient} is a type of orthogonal \gls{doe}, 
which has the advantage that each level for each variable will be tested only once. 
A Latin hypercube \gls{doe} can be also created such that data points more uniformly distribute over the search space. 
%"Our ability to design a good experiment should depend upon our prior knowledge regarding the nature of the data generating mechanism."\cite{goodwin1977experiment}
%
%Continuous state spaces must be accommodated by arbitrary discretization.
\cite{cohn1996neural}
%
\td{For a 2-level factorial applies: "The effect of any factor main effect (A, B, C) or interaction (AB, AC, BC, ABC) is the difference of two averages, the average of the responses correpsonding to +1 levels and the average of the responses corresponding to -1 levels. " - from Gunst2009\cite{gunst2009fractional}}

DOEs: randomization, latin square, orthogonal experiment design, full/2-lvl factorial design, placket burman, 
\url{https://www.jstor.org/stable/2238570}\cite{karlin1966optimal} very math intensive  \\
\url{https://doi.org/10.1111/j.2517-6161.1973.tb00944.x}\cite{whittle1973some}  \\
maybe \url{https://www.sciencedirect.com/science/article/pii/S0167715212003343?via\%3Dihub}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Machine Learning and Statistics}
\ds{
statistics:
	what is the difference between ml and stat? 
		superficially very similar but different goals and diff rahmenbedingungen 
    how can i usee ANOVA 
    linear regression 
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Why introduce machine learning to this project? 
%I wanted to apply and delve into what I've been studying during my undergrad courses. 
%This seemed like the perfect oppurtunity. 
Is machine learning just hyped statistics? No! 
Although machine learning uses several statistical methods, their goal and frame conditions are different. 
\td{The goal of machine learning is to ???}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Artificial Inteligence and Machine Learning}
%How can \gls{ai} and \gls{ml} be defined? 
The history of \gls{ai} goes back to the middle of the \nth{20} century. 
Researchers from the emerging field came together at the Dartmouth conference and the term "\gls{ai}" was coined\cite{McCarthy1955}. 
\Gls{ai}'s  history is beautifully depicted by McCorducks' 1982 book "Machines Who Think"\cite{McCorduck1982,Apter1982}, which focuses on the great minds behind the advances.
Pioneers like Alan Turing thought a lot about how to define, test and implement \gls{ai}\cite{Howard2019}. 
One example how to measure \gls{ai} is to let it play chess against a human\cite{Silver2017} 
(in 1997 a chess computer called Deep Blue won against the World Chess Champion Garry Kasparov for the first time\cite{Feng1999}).
Another test ingenioused by Alan Turing, is that a human communicates with an unknown entity (in written form) and must judge if they are dealing with a machine or a human being. 
It is easy for humans to come up with question to detect an AI, 
but when reading AI written articles\cite{gpt2020}, it's easy to see this test being passed in the near future.

@@

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Machine Learning Methods}
%
\Gls{ml} is at the base of most \gls{ai}s.
It is an umbrella term for programs with instructions to learn from data, i.e. gain knowledge, categorise, predict and make decisions based on data. 
%
There is a platitude of different machine learning methods: 
they can be divided into supervised (training set is labeled) and unsupervised (exploratory).  
An orthogonal division can be made by regression (continuous data) versus classification (discrete, categorical data). 
Independent from these $2\cdot2$ categories there are multiple ways to let machines learn from data.
%
\Gls{nn} (one of the most popular architectures for big data\cite{Chiroma2019}) are loosely modeled after the brain\cite{bishop1994neural}.
Artificial neurons (also called nodes), which are arranged in layers, 
are connected to each of the neurons of previous and next layers
and the weights (parameters of intensity), with which the data is routed from one neuron to another, 
are optimised during training. 
%
Convolutional-layer \gls{nn} excel in picture recognition\cite{Lecun1995conv} and in quantum mechanics too\cite{westermayr2020combining}.
Other common architectures include linear regression, kernel ridge regression and support vector regression.
There are also lesser know algorithms like 
evolutionary algorithms (e.g. \gls{ga} and \gls{pso}).
%
One advantage of evolutionary algorithms is that they start with a small data set
and periodically request new data in order to solve the problem iteratively.
\td{This is exactly what I needed because every experiment is very time intensive.}

@@

\Gls{pso} also uses a starting population of particles where each experiment (particle) 
is represented by its independent and dependent variables. 
It was originally inspired by the behaviour of bird flocks and fish schools\cite{Kennedy1995}.
Each particle has an associated position and velocity. 
Every movement across the search space is stochastic and influenced by its particle velocity and position as well as its and the swarm's best known position.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Design of Experiment} %DOE
%\epigraph{"The real purpose of experiment design is to maximize the information content of the data within the limits imposed by the given constraints."}{Grahem C. Goodwin\cite{goodwin1977experiment}}
%
\begin{quote}
	{"The real purpose of experiment design is to maximize the information content of the data within the limits imposed by the given constraints."}
	- {Grahem C. Goodwin\cite{goodwin1977experiment}}
\end{quote}
%
In two cases a deliberate \gls{doe} is beneficial.
If the query of a new data point is very expensive, it is favourable to actively chose the query (e.g. drilling for oil or quantum chemical calculations). 
If the query space is so vast, that randomly querying might explore domains which might lead to uninteresting or even misleading information.
% 
At the beginning of any experiment its constraints must be determined. 
%
Constraints for a given experiment include range of input and output variables as well as total time available and total number of samples/experiments that can be taken.\cite{goodwin1977experiment}
%

%FULL FACTORIAL
A naive approach to an experiment design is the full factorial \gls{doe}.
Each possible combination of discrete values is tested. 
While this is the most informative design it will more often than not be infeasible due the curse of dimensionality\cite{cherkassky1998learning}.
% 2-LVL FACTORIAL
The 2-level factorial design provides an alternative with $2^d$ (where $d$ is the number of independent variables) experiments. 
Drawbacks of 2-level factorial \gls{doe}s include no data about the inside of the search space and infeasibility for high dimensional problems.
In a full factorial or 2-level factorial design most experiments are redundant and most resources will be spent exploring high-order interaction effects\cite{gunst2009fractional}, which are often minimal to non-existent.
% RANDOM 
In order to overcome these obstacles a certain number of experiments can be chosen randomly from the search space. 
%
%A straight forward design can be to randomly chose a certain number of experiments from the search space. 
%Alternatively, a subset of from the 2-level factorial design can be chosen, which is called a fractional factorial design. 
When a subset is chosen from the factorial design, it is called a fractional factorial design. 
% PLACKETT_BURMAN
The Plackett-Burman\cite{vanaja2007design,miller2001using,wang1995hidden} design is a special case of 2-level fractional factorial design, 
where the number of needed experiments $n$ is $n<d+4$ 
(more precisely $n=(\lfloor d\div4\rfloor+1)\cdot4$, where $\lfloor x\rfloor$ denotes the floor function on $x$) 
and each combination of levels for any pair of factors appears the same number of times. 
% HAMMERSLEY 
A drawback of 2-level factorial and random fractional designs is that the sample set is likely not evenly distributed across the search space\cite{viana2016tutorial}. 
The Hammersley design\cite{viana2016tutorial,diwekar1997efficient} is based on the Hammersley sequence and produces space filling data points. 
% LATIN HYPER CUBE
The Latin hypercube \gls{doe}\cite{viana2016tutorial,diwekar1997efficient} is a type of orthogonal \gls{doe}, 
which has the advantage that each level for each variable will be tested only once. 
A Latin hypercube \gls{doe} can be also created such that data points more uniformly distribute over the search space. 
Latin hypercube \gls{doe}s are mainly used in computer simulations which are purely deterministic and therefore have no error. 
%"Our ability to design a good experiment should depend upon our prior knowledge regarding the nature of the data generating mechanism."\cite{goodwin1977experiment}
%
%Continuous state spaces must be accommodated by arbitrary discretization.
\cite{cohn1996neural}
%
\td{For any 2-level factorial DOE applies: "The effect of any factor main effect (A, B, C) or interaction (AB, AC, BC, ABC) is the difference of two averages, the average of the responses correpsonding to +1 levels and the average of the responses corresponding to -1 levels. " - from Gunst2009\cite{gunst2009fractional}}

DOEs: randomization, latin square, orthogonal experiment design, full/2-lvl factorial design, placket burman, \\
\url{https://doi.org/10.1111/j.2517-6161.1973.tb00944.x}\cite{whittle1973some}  \\
maybe \url{https://www.sciencedirect.com/science/article/pii/S0167715212003343?via\%3Dihub}
